Our initial algorithm embedded the search procedure into the PBS algorithm, however we noted PBS uses an non-standard genetic selection method where the population is updated as soon as a better solution is found (this is somewhat similar to elitism selection). The impact is the diversity of the population may be reduced, therefore we also tested roulette and tournament selection. We generated a small training set of 10 problems using the procedure described in \cref{alg:generate_colourful} and conducted 50 trials setting a time limit of $0.15n+0.5k$ seconds (our full methodology is described in \cref{section:methodology}). Our results are reported in \cref{table:compare_genetic}.

\input{2_body/background/compare_genetic_table}

All three methods achieved near optimal costs for train 1-7, however they all performed poorly on the larger instances train 8-10. In all intances the original PBS selection method performed equal to or better than the other two. To determine whether these results are statistically significant we performed a Quade test on the mean cost for each instance and algorithm; a statistically significant result would have a p-value below $\alpha =0.05$. Our resulting p-value is 0.02, showing the three methods are statistically significant. To determine which algorithms are statistically significant from eachother, we perform a post-hoc pairwise Quade test shown in \cref{table:compare_genetic_quade}.

\input{2_body/background/compare_genetic_post_hoc_quade}

The original PBS selection method is statistically significant and better than roulette selection, since $0.0225<0.05$ and it produced better results in 7/10 instances. Original PBS is also statistically significant from tournament selection although to a lesser degree since it falls under the 90\% confidence interval as $0.0795<0.1$, on average it produced results with lower cost for 7/10 instances. The differences between tournament and roulette selection are insignificant as $0.4373>0.1$. Therefore our final algorithm will use the original PBS selection method.