The robust $k$-center problem (also known as the $k$-center problem with outliers) extends $k$-center problem by introducing a constraint \(p\) (\(1 \leq p \leq n \)) which specifies the minimum number of clients that must be serviced the centers (\cite{charikar_algorithms_2001}). This problem is more applicable to real scenarios, in our parcel delivery service analogy (\cref{section:k_center}), there may be houses on the outskirts of the city which are not profitable to deliver to (since the clients are less concentrated in that area); therefore we may want to exclude them when building the warehouses. The city can change their requirements and ask to build $k$ warehouses such that 80\% of the city's population lives in a reasonable distance away from their nearest warehouse. We now have the robust $k$-center problem. Note that when \(p=n\), we have the $k$-center problem.

Formally, this problem can be stated as follows:

\textbox{Given a graph $G=(V, E)$ on a set of $n$ vertices $V = \{v_1,..., v_n\}$, a maximum number of centers $k$ and a minimum coverage constraint $p$ (where $1\leq k\leq n$ and $1\leq p\leq n$):

\vspace{0.2cm}We want to find a subset of centers $C\subseteq V$ which covers a subset of vertices $Z\subseteq V$ such that $|C|\leq k$ and $|Z|\geq p$ which minimises the objective function: $max_{v\in Z}min_{c\in C}dist(v, c)$}

\citeauthor{charikar_algorithms_2001} proposed a 3-approximation algorithm in $\mathcal{O}(n^{3})$ time. They note that the cost of the solution must be one of the $n^{2}$ interpoint distances in $\{d(i,j)\mid i\in V, j\in V\}$. Their method performs a binary search on the interpoint distances, at each iteration evaluating a cost $r$; given a radius $r$, a center which covers the most uncovered points within distance $r$ is chosen. The radius is expanded and all points within $3r$ are marked as covered. This is repeated until $k$ centers are chosen. The search looks for the smallest radius $r$ which covers more than $p$ points. For a more detailed description of this algorithm we refer the reader the original paper by \textcite{charikar_algorithms_2001} and a later paper which discussed implementation optimisations by \textcite{schwartz_efficient_2010}.